Counting infinitely by Oritatami co-transcriptional folding

### 2019-08-02 paper.tex
writting proof abstract

\begin{abstract}
abstract
\end{abstract}

%------------------------------------------------------
	\section{Introduction}
%------------------------------------------------------
Our body consists many proteins which is based on genetic sequence in DNA.
DNA sequence has to be copied in order to make proteins because DNA is in cell nucleus.
It is copied to RNA (and its sequence is called \textit{transcript}) by a RNA polymerase and delivered to out of nucleus.
A transcript consists four types nucleobases A, U, G, and C, and a nucleobase are called \textit{bead}
Moreover, they form hydrogen bonds mainly between A and U, and between G and C, then they begin to form at the same time synthesizing.
Once the bead is fixed after that it can not change the position.
By repeating this method, transcript form a 3D shape.

A shape of RNA depends on the DNA sequence.
Geary, Rothemund, and Andrsen have succeeded in experimentally making 2D RNA tiles (Fig.\ref{fig:rna_origami}) by repeatedly copying a specific DNA sequence \cite{GearyRothemundAndersen2014}.
Then, Oritatami System was designed as a mathematical model that the process of RNA production \cite{counter2016} and Fig.\ref{fig:origami_on_oritatami} is the 2D RNA tile which is represented on Oritatami System.
Oritatami system treats transcript and nucleobase as paths and vertices which called \textit{bead}, respectively, moreover, they are arranged on a triangular grid.
The system takes transcript as an input and output an terminal conformation.
Geary et al. implemented binary counter in this system \cite{counter2016}.
This counter is folded in zigzags, and each row represents a counted number.
Then, this counter is given a bit width as an input to the system before transfer begins, and encounters nondeterminism if it carries beyond that bit width.
Since it is running on deterministic system, it halts when it becomes nondeterministic.

In this paper, based on the implementation method of this counter, we implement a counter which defined the behavior that extends the bit width when it overflows.

\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{fig/rna_origami.pdf}
\caption{
A generated and stabilized 2D RNA tile from DNA
}
\label{fig:rna_origami}
\end{figure}

\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{fig/origami_oritatami.pdf}
\caption{
The represented 2D RNA tile on Oritatami System
}
\label{fig:origami_on_oritatami}
\end{figure}

@@

\subsection{Conformations of Module H}
The half adder module H forms five different paths as shown in Fig.~\ref{fig:halfadders}.
Module H reads the already folded bead (F0 or F1) at above as one of the inputs, and determines the other input depending on the folding starts (top or bottom).
Module H forms four bricks of H00, H01, H10, and H11 for each input.
Then, the bricks outputs the sum as beads of bottom row and also outputs carry depending on whether module H is folded top or bottom except He1.
Brick He1 is the first folded module H after overflow.
Since this counter uses three row of gliders as a self-standing method, module H consisting of 30 beads has an even width, and when it starts folding from the bottom, folding ends at the bottom.
That is why brick He1 ends folding at the bottom and outputs carry equal to 0.

@@

@@

>>>>

\begin{abstract}
A fixed bit-width counter was proposed as a proof-of-concept demonstration of an oritatami model of cotranscriptional folding [Geary et al., Proc.~MFCS 2016, LIPIcs 58, 43:1-43:14], and it was embedded into another oritatami system that self-assembles a finite portion of Heighway dragon fractal. 
In order to expand its applications, we endow this counter with capability to widen bit-width at every encounter with overflow. 
\end{abstract}

%------------------------------------------------------
	\section{Introduction}
%------------------------------------------------------

Counting is one of the most essential tasks for computing; as well known, it suffices to count for Turing universality \cite{Minsky1967}. 
Nature has been counting billions of days using molecular ``circadian clockwork'' which is ``as complicated and as beautiful as the wonderful chronometers developed in the 18th century''  \cite{McClung2006}. 
Nowadays, developments in molecular self-assembly technology enable us to design molecules to count. 
Evans has recently demonstrated a DNA tile self-assembly system that counts accurately in binary from a programmed initial count until it overflows \cite{EvansPhD}. 
In its foundational theory of molecular self-assembly, such binary counters have proved their versatility, being used to assemble shapes of particular size \cite{AdChGoHu2001,RothemundWinfree2000}, towards self-assembly of fractals \cite{MasudaSekiUbukata2018}, as an infinite scaffold to simulate all Turing machines in parallel in order to prove undecidability of nondeterminism in the abstract tile-assembly model \cite{BrChDoKaSe2013}, and so on.

\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{fig/rna_origami.pdf}
\caption{RNA origami}
\label{fig:rna_origami}
\end{figure}

A fixed bit-width (finite) binary counter has been implemented as a proof-of-concept demonstration of oritatami model of cotranscriptional folding \cite{GeMeScSe2019}. 
As shown in Fig.~\ref{fig:rna_origami}, an RNA transcript folds upon itself while being transcribed (synthesized) from its corresponding DNA template strand. 
Geary et al. programmed a specific RNA rectangular tile structure into a DNA template in such a way that the corresponding RNA transcript \textit{folds cotranscriptionally} into the programmed tile structure highly likely \textit{in vitro} at room temperatures (\textit{RNA origami}) \cite{GearyRothemundAndersen2014}. 
An oritatami system folds a transcript of abstract molecules called \textit{beads} of finitely many types cotranscriptionally over the 2-dimensional triangular lattice cotranscriptionally according to a rule set that specifies which types of molecules are allowed to bind. 
The transcript of the binary counter in \cite{GeMeScSe2019} is of period 60 as $\textcircled{\scriptsize 0}{-}\textcircled{\scriptsize 1}{-}\textcircled{\scriptsize 2}{-} \cdots {-}\textcircled{\scriptsize 58}{-}\textcircled{\scriptsize 59}{-}\textcircled{\scriptsize 0}{-}\textcircled{\scriptsize 1} \cdots$ and its period is semantically divided into two half-adder (HA) modules $A = \textcircled{\scriptsize 0}{-}\textcircled{\scriptsize 1}{-} \cdots {-}\textcircled{\scriptsize 11}$ and $C = \textcircled{\scriptsize 30}{-}\textcircled{\scriptsize 31}{-} \cdots {-}\textcircled{\scriptsize 41}$ and two structural modules $B$ and $D$, which are sandwiched by half-adder modules.
While being folded cotranscriptionally in zigzags, HA modules increment the current count $i$ by 1, which is initialized on a linear \textit{seed} structure, alike the Evans' counter, whereas structural modules $B$ and $D$ align HA modules properly and also make a turn at an end of the count $i$; $B$ guides the transcript from a zig to a zag ($\hookrightarrow$) while $D$ does from a zag to a zig ($\hookleftarrow$). 
This counter was embedded as a component of an oritatami system to self-assemble an arbitrary finite portion of Heighway dragon fractal \cite{MasudaSekiUbukata2018}. 
Its applications is limited, however, by lack of mechanism to widen bit-width; its behavior is undefined when its count overflows. 
In this paper, we endow this counter, or more precisely, its structural module $B$, with capability to widen the count by 1 bit at every encounter with overflow.

%Our body consists many proteins which is based on genetic sequence in DNA.
%DNA sequence has to be copied in order to make proteins because DNA is in cell nucleus.
%It is copied to RNA (and its sequence is called \textit{transcript}) by a RNA polymerase and delivered to out of nucleus.
%A transcript consists four types nucleobases A, U, G, and C, and a nucleobase are called \textit{bead}
%Moreover, they form hydrogen bonds mainly between A and U, and between G and C, then they begin to form at the same time synthesizing.
%Once the bead is fixed after that it can not change the position.
%By repeating this method, transcript form a 3D shape.
%
%A shape of RNA depends on the DNA sequence.
%Geary, Rothemund, and Andrsen have succeeded in experimentally making 2D RNA tiles (Fig.\ref{fig:rna_origami}) by repeatedly copying a specific DNA sequence \cite{GearyRothemundAndersen2014}.
%Then, Oritatami System was designed as a mathematical model that the process of RNA production \cite{counter2016} and Fig.\ref{fig:origami_on_oritatami} is the 2D RNA tile which is represented on Oritatami System.
%Oritatami system treats transcript and nucleobase as paths and vertices which called \textit{bead}, respectively, moreover, they are arranged on a triangular grid.
%The system takes transcript as an input and output an terminal conformation.
%Geary et al. implemented binary counter in this system \cite{counter2016}.
%This counter is folded in zigzags, and each row represents a counted number.
%Then, this counter is given a bit width as an input to the system before transfer begins, and encounters nondeterminism if it carries beyond that bit width.
%Since it is running on deterministic system, it halts when it becomes nondeterministic.
%
%In this paper, based on the implementation method of this counter, we implement a counter which defined the behavior that extends the bit width when it overflows.

%\begin{figure}[tb]
%\centering
%\includegraphics[width=\linewidth]{fig/rna_origami.pdf}
%\caption{
%A generated and stabilized 2D RNA tile from DNA
%}
%\label{fig:rna_origami}
%\end{figure}
%
%\begin{figure}[tb]
%\centering
%\includegraphics[width=\linewidth]{fig/origami_oritatami.pdf}
%\caption{
%The represented 2D RNA tile on Oritatami System
%}
%\label{fig:origami_on_oritatami}
%\end{figure}

@@

\subsection{Conformations of Module H}
The half adder module H forms five different paths as shown in Fig.~\ref{fig:halfadders} in zig pass.
Module H reads the already folded bead (F0 or F1) at above as one of the inputs, and determines the other input depending on the folding starts (top or bottom).
Module H forms four bricks of H00, H01, H10, and H11 for each input.
Then, the bricks outputs the sum as beads of bottom row and also outputs carry depending on whether module H is folded top or bottom except He1.
Brick He1 is the first folded module H after overflow.
Since this counter uses three row of gliders as a self-standing method, module H consisting of 30 beads has an even width, and when it starts folding from the bottom, folding ends at the bottom.
That is why brick He1 ends folding at the bottom and outputs carry equal to 0.

@@

\subsection{Conformations of Module L}
The left turn module L forms five different paths as shown in Fig~\ref{fig:leftturns}.
In normal zig zag folding, module L is a glider like Lt or Lbc, but when transcript of L binding to Turn Signal, it formed as a turn module like Ltrc or Ltre.
Module L in the form as turn module has Turn Signal in its ends and it helps other L turn at the next zig-zag.
In zig pass, when module L reaches the left end, if it does not overflow, it bind to this Turn Signal and turned.
If module L overflows, it forms Lbe and continues the zig pass.
Then, module L which is folded next to Lbe, forms Ltre and turns by binding to the top of module F that is just before it.
At this time, the top of module F functions as Turn Signal, but this is the behavior only overflow because otherwise there is some folded module above module F.

\begin{figure}[tb]
\begin{tabular}{c}
 \begin{minipage}{0.5\linewidth}
  \centering
   \includegraphics[width=0.8\linewidth]{fig/svg/Lt_1.pdf}
 \end{minipage}

 \begin{minipage}{0.5\linewidth}
  \centering
   \includegraphics[width=0.8\linewidth]{fig/svg/Lbc_1.pdf}

\ \\
\ \\
  \centering
   \includegraphics[width=0.4\linewidth]{fig/svg/Ltrc_1.pdf}

\begin{tabular}{c}
\ \\
\ \\
 \begin{minipage}{0.5\linewidth}
  \centering
   \includegraphics[width=0.9\linewidth]{fig/svg/Lbe_1.pdf}
 \end{minipage}

  \begin{minipage}{0.5\linewidth}
  \centering
   \includegraphics[width=0.9\linewidth]{fig/svg/Ltre_1.pdf}

 \end{minipage}
 \end{tabular}
 \caption{The paths of module L}
 \label{fig:leftturns}
\end{figure}

@@

\centering
\begin{table}[tb]
\begin{tabular}{c c c c}
\begin{minipage}{25mm}
	\centering
       \scalebox{0.5}{\includegraphics{fig/svg/bglider.pdf}}
\end{minipage} &
\begin{minipage}{25mm}
	\centering
       \scalebox{0.5}{\includegraphics{fig/svg/tglider.pdf}}
\end{minipage} &
\begin{minipage}{30mm}
	\centering
       \scalebox{0.5}{\includegraphics{fig/svg/plane.pdf}}
\end{minipage} &
\begin{minipage}{25mm}
	\centering
       \scalebox{0.5}{\includegraphics{fig/svg/spiral.pdf}}
\end{minipage}\\

\begin{minipage}{25mm}
	\centering
       \scalebox{0.5}{\includegraphics{fig/svg/horn.pdf}}
\end{minipage} &
\begin{minipage}{30mm}
	\centering
       \scalebox{0.5}{\includegraphics{fig/svg/kar.pdf}}
\end{minipage} &
\begin{minipage}{25mm}
	\centering
       \scalebox{0.5}{\includegraphics{fig/svg/roof.pdf}}
\end{minipage}\\

\begin{minipage}{25mm}
	\centering
       \scalebox{0.5}{\includegraphics{fig/svg/fault.pdf}}
\end{minipage} &
\begin{minipage}{30mm}
	\centering
       \scalebox{0.5}{\includegraphics{fig/svg/cliff.pdf}}
\end{minipage}\\

\end{tabular}
\end{table}

\begin{thebibliography}{1}

\bibitem{AdChGoHu2001}
Leonard Adleman, Qi Chang, Ashish Goel, and Ming-Deh Huang, 
Running time and program size for self-assembled squares, 
In Proc. STOC 2001, ACM, 740-748, 2001.

\bibitem{BrChDoKaSe2013}
Nathaniel Bryans, Ehsan Chiniforooshan, David Doty, Lila Kari, and Shinnosuke Seki, 
The power of nondeterminism in self-assembly, 
\textit{Theory of Computing}, vol. 9, 1-29, 2013.

\bibitem{EvansPhD}
Constantine Glen Evans, 
Crystals that Count! Physical Principles and Experimental Investigations of {DNA} Tile Self-Assembly, 
Ph.D. thesis, Caltech, 2014.

\bibitem{GeMeScSe2019}
Cody Geary, Pierre-\'{E}tienne Meunier, Nicolas Schabanel, and Shinnosuke Seki, 
Oritatami: A computational model for molecular co-transcriptional folding, 
\textit{International Journal of Molecular Sciences} vol. 20(9), 2259, 2019. Its conference version was published in Proc. MFCS 2016.

\bibitem{GearyRothemundAndersen2014}
Cody Geary, Paul W. K. Rothemund, and Ebbe S. Andersen, 
A single-stranded architecture for cotranscriptional folding of {RNA} nanostructures, 
\textit{Science} vol 345(6198), 799-804, 2014.

\bibitem{MasudaSekiUbukata2018}
Yusei Masuda, Shinnosuke Seki, and Yuki Ubukata, 
Towards the algorithmic molecular self-assembly of fractals by cotranscriptional folding, 
In Proc. CIAA 2018, LNCS 10977, Springer, 261-273, 2018.

\bibitem{McClung2006} 
C. Robertson McClung, 
Plant circadian rhythms, 
\textit{The Plant Cell}, vol. 18, 792-803, 2006.

\bibitem{Minsky1967}
Marvin Minsky, 
Computation: Finite and Infinite Machines, 
Prentice-Hall, Inc., 1967.

\bibitem{RothemundWinfree2000}
Paul W. K. Rothemund and Erik Winfree, 
The program-size complexity of self-assembled squares (extended abstract), 
In Proc. STOC 2000, ACM, 459-468, 2000.

\bibitem{WinfreePhD}
Erik Winfree, 
Algorithmic Self-Assembly of {DNA}, 
Ph.D. thesis, Caltech, 1998.

\end{thebibliography}
